### 2022-01-26 documentos/generators.tex
etc

Given that both matrix $\san$ and PCE generators $\pceg$ have the same 
symmetries, it is worth pointing out that $\san$ (and thus $a$) [see \eref{eq:san}] 
encodes all the information of PCE generators $\pceg$ and, therefore, of all PCE quantum channels. 
From the tensor power of matrix $a$ one can infer the $\{\taus\}$ of a PCE
generator $\pceg$ by taking any row or column $A^{(i)}$ and mapping 
all entries with $-1$ to $0$. In \Fref{fig:A_and_PCEGs_connection} we illustrate
this connection between $a$ and single qubit PCE generators.
In \appref{sec:2_qubits} we illustrate this program for 2-qubits PCE channels.

>>>>

\ddnote{Tengo entendido que esto se queria probar usando las simetrias, esto ya esta probado en la secc V}
Given that both matrix $\san$ and PCE generators $\pceg$ have the same 
symmetries, it is worth pointing out that $\san$ (and thus $a$) [see \eref{eq:san}] 
encodes all the information of PCE generators $\pceg$ and, therefore, of all PCE quantum channels. 
From the tensor power of matrix $a$ one can infer the $\{\taus\}$ of a PCE
generator $\pceg$ by taking any row or column $A^{(i)}$ and mapping 
all entries with $-1$ to $0$. In \Fref{fig:A_and_PCEGs_connection} we illustrate
this connection between $a$ and single qubit PCE generators.
In \appref{sec:2_qubits} we illustrate this program for 2-qubits PCE channels.

### 2022-01-26 documentos/kraus.tex
etc

% \cpnote{We might have spoken about the notation of the generators and here we 
% can add a sentence about it, but I will wait till JA writes his part.}
% Two remarks. (i) The notation that we develop with Jose Alfredo coincides with the 
% generator. This notation is based on the action on each individual qubit. (ii) Counting the
% number of generators here is trivial. It is simply $4^n-1$ as previously noticed 
% from other arguments. 
% }}}
\subsection{Pure dissipative implementation} % {{{
Any PCE channel can be seen as the fixed point of some decoherence process\cpnote{esto suena obvio, pero no lo es. Es decir no todos los canales son resultado de
lindblad. Me gustaria modificar esta frase para que pueda ser apreciada mejor}. We
will show this first for PCEGs. Consider the following dynamical process that
\textit{implements} \cpnote{porque en italicas? Me cae gordo} the PCEG $\mcG_{\vec \alpha}$ when $t\to \infty$,
\begin{align}
\mcG_{\vec \alpha,t}[\rho]&=e^{- \gamma t} \rho + (1-e^{-\gamma t})\mcG_{\vec \alpha}[\rho]\nonumber \\
&=\frac{\left( 1+e^{- \gamma t} \right)}{2} \rho +\frac{\left( 1-e^{- \gamma t} \right)}{2} \sigma_{\vec \alpha} \rho \sigma_{\vec \alpha},
\end{align}
where $\gamma>0$. It is easy to show that the family of channels $\mcG_{\vec
\alpha, t}$ parametrized with $t \geq 0$ forms a one-parametric semigroup,
\ie{} $\mcG_{\vec \alpha, t_1}\mcG_{\vec \alpha, t_2}=\mcG_{\vec \alpha,
t_1+t_2}$, therefore $\mcG_{\vec \alpha, t}$ describes a dissipative
time-homogeneous Markovian process, those are always characterized by some
Lindblad generator~\cite{lindblad}. The Lindblad generator of $\mcG_{\vec
\alpha, t}$, denoted by $\mcL_{\vec \alpha}$, can be obtained with the standard
procedure,  
\begin{equation}
\mcL_{\vec \alpha}[\rho]=\left. \frac{d\mcG_{\vec \alpha,t}[\rho]}{dt}\right|_{t=0}=\frac{\gamma\left(\sigma_{\vec \alpha}\rho \sigma_{\vec \alpha} -\rho \right)}{2},
\end{equation}
where the unique Lindblad operator associated with the relaxation ratio
$\gamma/2$ is simply $\sigma_{\vec \alpha}$. Notice that $\sigma_{\vec \alpha}$
is trace-less, therefore the process is pure dissipative~\cite{cirac}.

Since PCEGs commute, we can describe easily any other PCE channel as a fixed
point of a decoherence process. For them, the Lindblad generators are the sum
of the Lindbladians of the corresponding generators. As an example, consider
the channel depicted in \fref{fig:two:qubit:examples} b), it is equal to
$\mcG_{(0,3)}\mcG_{(3,3)}$, therefore it is the fixed point of the dissipation
process described with the following Lindbladian,
\begin{equation}
\mcL=\frac{\gamma_{(0,3)}\left(\sigma_{(0,3)}\rho\sigma_{(0,3)}-\rho \right)}{2}
   +\frac{\gamma_{(3,3)}\left(\sigma_{(3,3)}\rho\sigma_{(3,3)}-\rho
\right)}{2},
\end{equation}
\cpnote{Aca cambio la notacion, no se deberia poner L[rho] ?}
where $\gamma_{(0,3)}$ and $\gamma_{(3,3)}$ are positive and correspond to the Lindblad operators $\sigma_{(0,3)}$ and $\sigma_{(3,3)}$. Notice that such election of Lindblad operators is not unique, as the PCE channel described here is also equal to $\mcG_{(0,3)}\mcG_{(3,0)}$.

% }}}
\subsection{Collision model implementation} % {{{
We now show that PCE channels can also be described with \textit{simple
collision models}~\cite{Ziman2010} \cpnote{can also be implemented can result 
from collision models o ago asi. No cre oqeu uno describa un PCE con un modelo
de colision}. First, observe that PCEGs can
be implemented using an unitary over the system and an
ancilla~\cite{Stinespring2006}\cpnote{No creo qeu la cita vaya acá}. Since PCEG
always have Kraus rank $2$, one can
always choose a qubit as the ancillary system. Concretely,
\begin{equation}
\mcG_{\vec \alpha}[\rho]=
    \tr_{\text{qubit}} \left( 
         U_{\vec \alpha} \left(\rho\otimes \projj{0}\right) U^\dagger{}_{\vec \alpha} 
     \right),
\end{equation}
where $\tr_{\text{qubit}}$ denotes partial trace over the ancillary qubit, with
the unitary defined as follows,
\begin{align}
U_{\vec\alpha}\ket{\psi} \ket{0}= \frac{1}{\sqrt{2}} 
   \left( \ket{\psi} \ket{0}+\sigma_{\vec\alpha} \ket{\psi}\ket{1} \right),\\
U_{\vec\alpha}\ket{\psi} \ket{1}=\frac{1}{\sqrt{2}} 
   \left( \ket{\psi} \ket{0}-\sigma_{\vec\alpha} \ket{\psi}\ket{1} \right).
\label{eq:Nqubit_stinespring}
\end{align}
Therefore, any concatenation of PCEGs can be described as a collision model
with as many collisions as generators needed. In fact, generators are described
with one collision. For the general case consider some PCE channel $\mcE$
generated with $\left\{ \mcG_{\vec \alpha_1}, \mcG_{\vec \alpha_2}, \dots,
\mcG_{\vec \alpha_M} \right\}$. For this we can define an environment
consisting of $M$ qubits initially in the state $\left( \projj{0}
\right)^{\otimes M}$, or equivalently one qubit with the additional assumption
that its state is reset to $\ket{0}$ after each collision (memory-less
collisions). The collision with the $k\text{th}$ particle is described by
$U_{\vec \alpha_k}$, which acts solely over the system and the $k\text{th}$
particle. Therefore $\mcE$ can be written as follows,
\begin{equation}
\mcE[\rho]=\tr_{\text{E}} \left[ \left(U_{\vec \alpha_1} \dots U_{\vec \alpha_M}\right) \rho \otimes \left(\projj{0}\right)^{\otimes M} \left(U_{\vec \alpha_1} \dots U_{\vec \alpha_M} \right)^\dagger{} \right],
\end{equation}
where $\tr_{\text{E}}$ is the partial trace over the total of the ancillary qubits. Notice that as PCEGs commute, the order of the collisions can be any.
% \begin{itemize}
% \item Relate to decoherence, for PCEG\ddnote{hecho}
% \item Play with non PCEG PCEs\ddnote{hecho}
% \item Dependiendo de si sale algo bonito, podemos poner algun modelo en
% particular para todos los canales (sistema central + ambiente). Seria de buscar
% dilataciones sencillas y razonables por lo menos a los canales generadores.\ddnote{hecho, es el asunto de collision models}
% \todo{Davalos lidera la discusión de ese punto, pero no escribe nada}
% \end{itemize} 
% 
% \begin{equation}
% \mcL_{\vec \alpha}[\Delta]:= \frac{\gamma}{2}\left( \sigma_{\vec\alpha} \Delta \sigma_{\vec\alpha} -\Delta \right).
% \label{eq:Nqubit_semigroup}
% \end{equation}

>>>>

% \cpnote{We might have spoken about the notation of the generators and here we 
% can add a sentence about it, but I will wait till JA writes his part.}
% Two remarks. (i) The notation that we develop with Jose Alfredo coincides with the 
% generator. This notation is based on the action on each individual qubit. (ii) Counting the
% number of generators here is trivial. It is simply $4^n-1$ as previously noticed 
% from other arguments. 
% }}}
\subsection{Pure dissipative implementation} % {{{
Any PCE channel can be seen as the fixed point of some decoherence process\cpnote{esto suena obvio, pero no lo es. Es decir no todos los canales son resultado de
lindblad. Me gustaria modificar esta frase para que pueda ser apreciada mejor}. We
will show this first for PCEGs. Consider the following dynamical process that
implements \cpnote{porque en italicas? Me cae gordo} the PCEG $\mcG_{\vec \alpha}$ when $t\to \infty$,
\begin{align}
\mcG_{\vec \alpha,t}[\rho]&=e^{- \gamma t} \rho + (1-e^{-\gamma t})\mcG_{\vec \alpha}[\rho]\nonumber \\
&=\frac{\left( 1+e^{- \gamma t} \right)}{2} \rho +\frac{\left( 1-e^{- \gamma t} \right)}{2} \sigma_{\vec \alpha} \rho \sigma_{\vec \alpha},
\end{align}
where $\gamma>0$. It is easy to show that the family of channels $\mcG_{\vec
\alpha, t}$ parametrized with $t \geq 0$ forms a one-parametric semigroup,
\ie{} $\mcG_{\vec \alpha, t_1}\mcG_{\vec \alpha, t_2}=\mcG_{\vec \alpha,
t_1+t_2}$, therefore $\mcG_{\vec \alpha, t}$ describes a dissipative
time-homogeneous Markovian process, those are always characterized by some
Lindblad generator~\cite{lindblad}. The Lindblad generator of $\mcG_{\vec
\alpha, t}$, denoted by $\mcL_{\vec \alpha}$, can be obtained with the standard
procedure,  
\begin{equation}
\mcL_{\vec \alpha}[\rho]=\left. \frac{d\mcG_{\vec \alpha,t}[\rho]}{dt}\right|_{t=0}=\frac{\gamma\left(\sigma_{\vec \alpha}\rho \sigma_{\vec \alpha} -\rho \right)}{2},
\end{equation}
where the unique Lindblad operator associated with the relaxation ratio
$\gamma/2$ is simply $\sigma_{\vec \alpha}$. Notice that $\sigma_{\vec \alpha}$
is trace-less, therefore the process is pure dissipative~\cite{cirac}.

Since PCEGs commute, we can describe easily any other PCE channel as a fixed
point of a decoherence process. For them, the Lindblad generators are the sum
of the Lindbladians of the corresponding generators. As an example, consider
the channel depicted in \fref{fig:two:qubit:examples} b), it is equal to
$\mcG_{(0,3)}\mcG_{(3,3)}$, therefore it is the fixed point of the dissipation
process described with the following Lindbladian,
\begin{equation}
\mcL[\rho]=\frac{\gamma_{(0,3)}\left(\sigma_{(0,3)}\rho\sigma_{(0,3)}-\rho \right)}{2}
   +\frac{\gamma_{(3,3)}\left(\sigma_{(3,3)}\rho\sigma_{(3,3)}-\rho
\right)}{2},
\end{equation}
\cpnote{Aca cambio la notacion, no se deberia poner L[rho] ?}
where $\gamma_{(0,3)}$ and $\gamma_{(3,3)}$ are positive and correspond to the Lindblad operators $\sigma_{(0,3)}$ and $\sigma_{(3,3)}$. Notice that such election of Lindblad operators is not unique, as the PCE channel described here is also equal to $\mcG_{(0,3)}\mcG_{(3,0)}$.

% }}}
\subsection{Collision model implementation} % {{{
We now show that PCE channels can also be implemented with \textit{simple
collision models}~\cite{Ziman2010} \cpnote{can also be implemented can result 
from collision models o ago asi. No cre oqeu uno describa un PCE con un modelo
de colision}. First, observe that PCEGs can
be implemented using an unitary over the system and an
ancilla~\cite{Stinespring2006}\cpnote{No creo qeu la cita vaya acá}. Since PCEG
always have Kraus rank $2$, one can
always choose a qubit as the ancillary system. Concretely,
\begin{equation}
\mcG_{\vec \alpha}[\rho]=
    \tr_{\text{qubit}} \left( 
         U_{\vec \alpha} \left(\rho\otimes \projj{0}\right) U^\dagger{}_{\vec \alpha} 
     \right),
\end{equation}
where $\tr_{\text{qubit}}$ denotes partial trace over the ancillary qubit, with
the unitary defined as follows,
\begin{align}
U_{\vec\alpha}\ket{\psi} \ket{0}= \frac{1}{\sqrt{2}} 
   \left( \ket{\psi} \ket{0}+\sigma_{\vec\alpha} \ket{\psi}\ket{1} \right),\\
U_{\vec\alpha}\ket{\psi} \ket{1}=\frac{1}{\sqrt{2}} 
   \left( \ket{\psi} \ket{0}-\sigma_{\vec\alpha} \ket{\psi}\ket{1} \right).
\label{eq:Nqubit_stinespring}
\end{align}
Therefore, any concatenation of PCEGs can be described as a collision model
with as many collisions as generators needed. In fact, generators are described
with one collision. For the general case consider some PCE channel $\mcE$
generated with $\left\{ \mcG_{\vec \alpha_1}, \mcG_{\vec \alpha_2}, \dots,
\mcG_{\vec \alpha_M} \right\}$. For this we can define an environment
consisting of $M$ qubits initially in the state $\left( \projj{0}
\right)^{\otimes M}$, or equivalently one qubit with the additional assumption
that its state is reset to $\ket{0}$ after each collision (memory-less
collisions). The collision with the $k\text{th}$ particle is described by
$U_{\vec \alpha_k}$, which acts solely over the system and the $k\text{th}$
particle. Therefore $\mcE$ can be written as follows,
\begin{equation}
\mcE[\rho]=\tr_{\text{E}} \left[ \left(U_{\vec \alpha_1} \dots U_{\vec \alpha_M}\right) \rho \otimes \left(\projj{0}\right)^{\otimes M} \left(U_{\vec \alpha_1} \dots U_{\vec \alpha_M} \right)^\dagger{} \right],
\end{equation}
where $\tr_{\text{E}}$ is the partial trace over the total of the ancillary qubits. Notice that as PCEGs commute, the order of the collisions can be any.
% \begin{itemize}
% \item Relate to decoherence, for PCEG\ddnote{hecho}
% \item Play with non PCEG PCEs\ddnote{hecho}
% \item Dependiendo de si sale algo bonito, podemos poner algun modelo en
% particular para todos los canales (sistema central + ambiente). Seria de buscar
% dilataciones sencillas y razonables por lo menos a los canales generadores.\ddnote{hecho, es el asunto de collision models}
% \todo{Davalos lidera la discusión de ese punto, pero no escribe nada}
% \end{itemize} 
% 
% \begin{equation}
% \mcL_{\vec \alpha}[\Delta]:= \frac{\gamma}{2}\left( \sigma_{\vec\alpha} \Delta \sigma_{\vec\alpha} -\Delta \right).
% \label{eq:Nqubit_semigroup}
% \end{equation}

### 2022-01-26 documentos/quantum_classical.tex
etc

\section{Quantum-Classical channels}
\label{quantum_classical}
A Quantum-classical (QC) channel is defined using an orthonormal basis in the Hilbert space. Let $B=\left\{ \ket{\psi_i} \right\}$ such a basis in $\hilbert$ with $\dim(\hilbert)=2^N$, the QC associated to $B$ is
\begin{equation}
\mcE^{\text{QC}}_B[\rho]=\sum_{i=1}^{2^N} \bra{\psi_i}\rho \ket{\psi_i} \projj{\psi_i},
\end{equation}
this is, QC channels project density matrices onto the corresponding diagonal matrix in this basis~\cite{nathanson2007pauli}.

Consider now that the basis $B$ is the common eigenbasis of a maximal set of commuting Pauli strings denoted by $\text{set}(B)$, such set contains $2^N$ elements (including the identity in each set), and there are $2^N+1$ of such sets~\cite{Lawrence2002}. We will prove that the QCs defined in this way are \pce{} channels with $2^N$ on their diagonal.

First we compute the components of $\mcE^{\text{QC}}_B$ in the basis $1/2^{N/1}\left\{\sigma_{\vec \alpha}\right\}$,
\begin{equation}
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l}=\frac{1}{2^N}\sum_{i=1}^{2^N} \bra{\psi_i}\sigma_{\vec k} \ket{\psi_i}\bra{\psi_i}\sigma_{\vec l} \ket{\psi_i}.
\end{equation}
To evaluate it observe that $\bra{\psi_i}\sigma_{\vec k} \ket{\psi_i}^2=1 \ \ \forall \sigma_{\vec k}\in \text{set}(B)$, and from the formula for the purity of $\ket{\psi_i}$,
\begin{equation}
1=\frac{1}{2^N}\sum_{\vec k} \bra{\psi_i} \sigma_{\vec k} \ket{\psi_i}^2,
\end{equation}
it follows that $\bra{\psi_i} \sigma_{\vec k'} \ket{\psi_i}=0 \ \ \forall \sigma_{\vec k'} \not\in \text{set}(B)$ since there only positive terms in the sum, and $2^N$ of them are already equal to $1$. Therefore,
\begin{align}
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec k}&=1,\nonumber\\ \left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l'}&=\left(\mcE^{\text{QC}}_B\right)_{\vec l' \vec l'}=0 \ \ \forall \sigma_{\vec k} \in \text{set}(B) \ \ \forall \sigma_{\vec l'} \not\in \text{set}(B).
\end{align}
To compute $\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l}$ for $\sigma_{\vec k}, \sigma_{\vec l} \in \text{set}(B)$ with $\vec k \neq \vec l$, observe that $\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l}=\left(\mcE^{\text{QC}}_B\right)_{\vec l \vec k}$, \ie{} the matrix corresponding to $\mcE^{\text{QC}}$ is an orthogonal projector. Thus, consider the blocks,
\begin{equation}
\left[ \begin{array}{cc}
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec k} & \left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l} \\ 
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l} & \left(\mcE^{\text{QC}}_B\right)_{\vec l \vec l}
\end{array}  \right]=\left[ \begin{array}{cc}
1 & \left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l} \\ 
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l} & 1
\end{array}  \right],
\end{equation}
it is easy to check that the latter is a projector only if $\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l}=0$. Since there are exactly $2^N$ elements

>>>>

\section{Quantum-Classical channels}
\label{quantum_classical}
\ddnote{Alguien revise este apendice, tiene solo una iterada}
A Quantum-classical (QC) channel is defined using an orthonormal basis in the Hilbert space. Let $B=\left\{ \ket{\psi_i} \right\}$ such a basis in $\hilbert$ with $\dim(\hilbert)=2^N$, the QC associated to $B$ is
\begin{equation}
\mcE^{\text{QC}}_B[\rho]=\sum_{i=1}^{2^N} \bra{\psi_i}\rho \ket{\psi_i} \projj{\psi_i},
\end{equation}
this is, QC channels project density matrices onto the corresponding diagonal matrix in the basis $B$~\cite{nathanson2007pauli}.

Consider now that the basis $B$ is the simultaneous eigenbasis of a maximal set of commuting Pauli strings denoted by $\text{set}(B)$, such set contains $2^N$ elements (including the identity), and there are $2^N+1$ of such sets~\cite{Lawrence2002}. We will prove that the QCs defined in this way are \pce{} channels with $2^N$ on their diagonal.

First we compute the components of $\mcE^{\text{QC}}_B$ in the basis $2^{-N/2}\left\{\sigma_{\vec \alpha}\right\}$,
\begin{equation}
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l}=\frac{1}{2^N}\sum_{i=1}^{2^N} \bra{\psi_i}\sigma_{\vec k} \ket{\psi_i}\bra{\psi_i}\sigma_{\vec l} \ket{\psi_i}.
\end{equation}
To evaluate the components, observe that $\bra{\psi_i}\sigma_{\vec k} \ket{\psi_i}^2=1 \ \ \forall \sigma_{\vec k}\in \text{set}(B)$, and from the formula for the purity of $\ket{\psi_i}$,
\begin{equation}
1=\frac{1}{2^N}\sum_{\vec k} \bra{\psi_i} \sigma_{\vec k} \ket{\psi_i}^2,
\end{equation}
it follows that $\bra{\psi_i} \sigma_{\vec k'} \ket{\psi_i}=0 \ \ \forall \sigma_{\vec k'} \not\in \text{set}(B)$ since there only positive terms in the sum, and $2^N$ of them are already equal to $1$. Therefore,
\begin{align}
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec k}&=1,\nonumber\\ \left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l'}&=\left(\mcE^{\text{QC}}_B\right)_{\vec l' \vec l'}=0 \ \ \forall \sigma_{\vec k} \in \text{set}(B) \ \ \forall \sigma_{\vec l'} \not\in \text{set}(B).
\end{align}
To compute $\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l}$ for $\sigma_{\vec k}, \sigma_{\vec l} \in \text{set}(B)$ with $\vec k \neq \vec l$, observe that $\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l}=\left(\mcE^{\text{QC}}_B\right)_{\vec l \vec k}$, \ie{} the matrix corresponding to $\mcE^{\text{QC}}$ is an orthogonal projector. Thus, consider the block,
\begin{equation}
\left[ \begin{array}{cc}
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec k} & \left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l} \\ 
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l} & \left(\mcE^{\text{QC}}_B\right)_{\vec l \vec l}
\end{array}  \right]=\left[ \begin{array}{cc}
1 & \left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l} \\ 
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l} & 1
\end{array}  \right],
\end{equation}
it is easy to check that the latter is a projector only if $\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l}=0$. Since there are exactly $2^N$ elements of the form $\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec k}$ with $\sigma_{\vec k} \in \text{set}(B)$, the channel $\mcE^{\text{QC}}_B$ is \pce{} with $2^N$ $1$s on its diagonal.
